I would like to thank my main supervisor Richard Morris for all the support he has given me throughout these long, and sometimes trying, four years.
I appreciate his giving me the opportunity to study for a PhD in his group and for allowing me to attempt to combine my mathematical knowledge with my interest in plants.
He has been full of useful suggestions, especially when things did not turn out as expected in this project.
His good humour and scientific integrity have been sterling examples for me to follow in the future.
I also thank Richard a lot for comments on draft chapters of this thesis which have improved it considerably.

I would like to thank all the post-docs and students in the Computational \& Systems Biology department who have been a great group to spend my time with.
In particular I'd like to acknowledge the help of David Richards in the early days for getting me up and running with learning Linux and Python.
He also orchestrated much lunchtime entertainment which I think provide many of us with fond memories.
I owe a lot of thanks to Matthew Hartley who joined CSB as our lab manager and was able to give me lots of hints on computing and programming.

Other members of my supervisory team deserve mentions.
Phil Wigge and Katja Jaeger provided the experimental angle of the project and I thank them for the leaf number data, and helping choose me to do a PhD at JIC\@.
Rico Coen also provided food for thought throughout the project and allowed us the benefit of his wisdom.

I am indebted to the Brassica tendency crew of Marc Jones, Rachel Wells, Judith Irwin, Martin Trick and Colin Morgan for making the final year of my PhD the most enjoyable one.
We've had many fun meetings, social occasions and they have allowed me to finally enter the greenhouses to see growing plants close up.
I also thank Judith for some useful comments on chapter 3.

It was also good to meet lots of people on the various training courses or conferences I attended all over Britain and Europe.
Their enthusiasm showed me a good view of science and the wider world.
I should also acknowledge lots of online bloggers and communities whose various tips saved me (or sometimes caused me!) much frustration at the keyboard.
Principally those who have, and continue to contribute to \url{StackOverflow.com} or \url{tex.stackexchange.com} have been incredibly valuable to my development as someone who can now program a fair bit.

In Norwich, I thank Glen and Annika Richardson for letting me lodge in their lovely house for the first nine months of my PhD.
Thereafter I thank Steve Garrett for being my landlord in a friendly house very close to work.
To the members of the UEA Athletics and Hockey Clubs I was a member of during my PhD years I am indebted to helping me keep it real, and reminding me life isn't all about about long hours in front of the office computer.
It was an honour to serve as UEAHC Club Captain and to help coach the distance athletes from UEAAC\@.

I feel very lucky to have met Yan Ma who has shown much kindness and support to me.
As a dedicated and hard-working PhD student she has been a great example of carrying on when times get tough.
Her knowledge and enthusiasm have helped me to gain a far greater understanding of many aspects of biological experiments.
She also provided the Arabidopsis plants I took photos of in chapter~1 and helped me with the photo editing.

Lastly, I thank my family for always being there.
They have always encouraged me and seem much more convinced of my capability than I am.
I wish them all the happiness for the future.

\vspace{1cm}
\noindent Nick Pullen\newline
{\scshape John Innes Centre, Norwich}\newline
September {\scshape 2014}
