\documentclass{nicksclass}

\includeonly{%Easy include/exclude of chapters
abstract,%
intro%
,
nestedSampling%
,
flowering%
,
discussion%
}

\begin{document}
\DeclareFieldFormat[article,book,manual,inproceedings,incollection,thesis,online]{title}{\mkbibbold{#1}}% Bold titles in Bibliography
\DeclareFieldFormat[article]{titlecase}{\MakeSentenceCase{#1}}
\strictpagecheck% For ouside margins

% Title page
\thispagestyle{empty}
\newgeometry{textheight=706pt, marginparsep=0pt, marginparwidth=0pt, textwidth=400pt, inner=98.753pt}%So inner = outer ergo text can be centred on page
\myTitlePage
\restoregeometry
\thispagestyle{empty}

% Abstract
\cleardoublepage
\thispagestyle{empty}
\phantomsection% Needed for hyperref to link correctly
\addcontentsline{toc}{chapter}{Abstract}
\chapter*{Abstract}
\input{abstract}
\thispagestyle{empty}
% only clearpage gets rid of following header
\clearpage
\thispagestyle{empty}

% ToC
{\hsize 441pt
\tableofcontents
\cleardoublepage}

% My publications
\phantomsection% Needed for hyperref to link correctly
\addcontentsline{toc}{chapter}{List of publications}
\chapter*{List of Publications}
  This thesis includes material from the published works below.\newline
  * Indicates that these authors contributed equally.

\vspace{1cm}\noindent
A. Bentley, E. Jensen, I. Mackay, H. Hönicka, M. Fladung, K. Hori, M. Yano, J. Mullet, I. Armstead, C. Hayes, D. Thorogood, A. Lovatt, R. Morris, N. Pullen, E. Mutasa-Göttgens, and J. Cockram. \textbf{Flowering Time}. In: \emph{Genomics and Breeding for Climate-Resilient Crops: Vol. 2 Target Traits} (2013), 1--66.

\vspace{1cm}\noindent
K. E. Jaeger*, N. Pullen*, S. Lamzin, R. J. Morris, and P. A. Wigge. \textbf{Interlocking feedback loops govern the dynamic behavior of the floral transition in \emph{Arabidopsis}}. \emph{The Plant Cell} 25.3 (2013), 820--833

\vspace{1cm}\noindent
N. Pullen, K. E. Jaeger, P. A. Wigge, and R. J. Morris. \textbf{Simple network motifs can capture key characteristics of the floral transition in \emph{Arabidopsis}}. \emph{Plant Signaling \& Behavior} 8.11 (2013), e26149.

\vspace{1cm}\noindent
N. Pullen and R. J. Morris. \textbf{Bayesian model comparison and parameter inference in systems biology using nested sampling}. \emph{PLoS ONE} 9.2 (2014), e88419.
\thispagestyle{empty}

% only clearpage gets rid of following header
\clearpage
\thispagestyle{empty}
\chapter*{Acknowledgements}
\chaptermark{Acknowledgements}% So correct name appears in header
\addcontentsline{toc}{chapter}{Acknowledgements}
\input{acknowledgements}

% Main text
\include{intro}
\include{nestedSampling}
\include{flowering}
\include{discussion}

% Bibliography - make wider so uses fewer pages in end
% and set ragged as its fairly discontinuous
\thispagestyle{empty}
\cleardoublepage
\newgeometry{textheight=686pt, marginparsep=0pt, marginparwidth=0pt, inner=40mm, outer=15mm, headsep=40pt, headheight=20pt}
\setlength{\topmargin}{-36pt}
\raggedright
\pagestyle{biblio}
\cleardoublepage
\phantomsection
\addcontentsline{toc}{chapter}{Bibliography}
\printbibliography%

% Colophon
\restoregeometry% back to easy to read textwidth
\thispagestyle{empty}
\chapter*{Colophon}
\flushboth% back to full justification
\input{colophon}
\phantomsection
\addcontentsline{toc}{chapter}{Colophon}

\end{document}
